---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  config_general = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)

devtools::load_all()
```

```{r child="helpers.Rmd", include=FALSE}

```

The docker image [rocker/verse](https://hub.docker.com/r/rocker/verse) provides
most of what you need for a functional computing environment, including system
dependencies, a bash terminal and RStudio. The image
[2dii/pacta_analysis](https://hub.docker.com/r/2dii/pacta_analysis) extends
rocker/verse to include the required R packages. Here is its Dockerfile:

```{r echo=FALSE, comment=""}
look_into("Dockerfile")
```

Finally you need access to the PACTA [siblings](#Siblings). You can
create one or more volumes mapping the location of the siblings from your (host)
computer to the container. 

Here are two ways to access a full computing environment, including PACTA
siblings. The code below assumes you have this .env file (added to .gitignore)
at the root of the repository, but you may change the value `abc` to anything.

```bash
# .env
RSTUDIO_PASSWORD=abc
```

This starts RStudio at http://localhost:8787. Login as "rstudio" with the
password you chose.

```bash
docker-compose up
```

This starts an interactive terminal, mounting your home directory at the
container's `/root` directory.

```bash
docker run --rm -ti -v "$HOME:/root" -w /root/PACTA_analysis  2dii/pacta_analysis:latest /bin/bash
```

For an improved environment replace `/bin/bash` with `docker/entrypoint`.

Reference:

* [Install Docker](https://docs.docker.com/engine/install/).
* [Install docker-compose](https://docs.docker.com/compose/install/).

<details>
```{r}
devtools::session_info()
```
</details>
