---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PACTA.analysis

<!-- badges: start -->
<!-- badges: end -->

The goal of this repository is to assess how well a porfolio aligns with climate
goals. This document provides a reproducible example of what you can achieve
with the code in this repository and friends.

You may want to analyze a single portfolio and investor, or multiple ones. With
2DII readers in mind, this document details two workflows:

* __Single-inputs workflow__: When the number of portfolios and investors is
exactly one, we now use a specific set of optimized scripts. (Because this
workflow is available online at <https://platform.transitionmonitor.com/start>,
it has been referred to as "the online workflow", and the scripts as "the web
tool scripts". This is misleading because this workflow can run offline too.)
* __Multiple-outputs workflow__: When the number of portfolios and/or investors
is not exactly one, we use a different set of optimized scripts. (By contrast
with the so-called "online" workflow, this one has been usually referred to as
"the offline workflow"; again, this is misleading.)

Users outside 2DII may instead see other related work:

* <https://platform.transitionmonitor.com/start>
* <https://github.com/2DegreesInvesting/r2dii.data>
* <https://github.com/2DegreesInvesting/r2dii.match>
* <https://github.com/2DegreesInvesting/r2dii.analysis>

## Installation

Get the source code of this and related repositories (more on this below) from
https://github.com/2DegreesInvesting/. You may use `git clone <URL>`.

## Example

```{r}
# TODO
```


