---
output: github_document
date: "`r Sys.time()`."
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)

devtools::load_all()
```

# PACTA_analysis

The goal of this repository is to assess how well a porfolio aligns with climate
goals. 

This documents targets internal 2DII users and developers. It provides
reproducible examples of what you can achieve with the code in this repository
and friends. You may use it as a guide to run your own analyses, or as an
integration test to ensure code changes preserve the behavior documented here.
Readers outside 2DII may instead see other related work:
[transitionmonitor.com](https://platform.transitionmonitor.com/start),
[r2dii.data](https://github.com/2DegreesInvesting/r2dii.data),
[r2dii.match](https://github.com/2DegreesInvesting/r2dii.match), and
[r2dii.analysis](https://github.com/2DegreesInvesting/r2dii.analysis).

You may want to analyze a single portfolio and investor, or multiple ones. This document details two workflows:

* __Single-inputs workflow__: When the number of portfolios and investors is
exactly one, we use a specific set of scripts optimized for this purpose. This
workflow is typically used to generate interactive reports. Because this
workflow is available [online](https://platform.transitionmonitor.com/start), it
has been referred to as "the online workflow", and the scripts as "the web tool
scripts". This is misleading because this workflow can run offline too.
* __Multiple-outputs workflow__: When the number of portfolios and/or investors
is not exactly one, we use a different set of optimized scripts. This workflow
is typically used to produce static (.pdf) reports. By contrast with the
so-called "online" workflow, this one has been usually referred to as "the
offline workflow"; again, this is misleading.

Both workflows include three steps:

1. Clean portfolio: Clean the input portfolio and merge in financial data, to
categorize each holding and identify whether it's equity (EQ) or corporate bonds
(CB).
2. Run Analysis: Merge portfolios with asset-level data and scenarios, then
group results at company, portfolio and regional level.
3. Present results: Present the results in a clear output format.

--

Note that "`$`" and "`#>`" precede bash and R output, respectively.

### Siblings

PACTA_analysis depends on other "siblings" of the PACTA family. They are all
available at <https://github.com/2DegreesInvesting> and should live alongside
PACTA_analysis/, under the same parent directory. All siblings include an empty file called ".pacta", which you can use to find them programmatically.

```{bash comment="$"}
siblings="$(dirname $(find .. -maxdepth 2 -empty -type f -name '.pacta'))"
echo $siblings
```

Before any analysis, ensure all siblings are up to date. For an easier way to
work with all siblings at once from the terminal see
[pacta-cli](https://github.com/2DegreesInvesting/pacta-cli).

## Computing environment

A functional computing environment is available in [this docker
image](https://github.com/2DegreesInvesting/docker/tree/master/r-packages) (how
to [install Docker](https://docs.docker.com/engine/install/)).

<details> <summary>Session information.</summary>
```{r}
devtools::session_info()
```
</details>

As an example, this is one way to use that computing environment to produce the
very document you are now reading:

* I move to the parent directory of PACTA_analysis and its siblings -- which in
my case is "~/git/".

```bash
cd ~/git
```

* I run (`docker run`) an ephemeral (`--rm`), interactive (`-ti`) container of
the image `2dii/r-packages`; create a number of volumes (`-v`) to access useful
files in my host computer from inside the container; and set the working
directory (`-w`) to PACTA_analysis/.

```{r comment="", echo=FALSE}
url <- "https://raw.githubusercontent.com/maurolepore/bin/master/docker-run-2dii-r-packages"
tmp <- tempfile()
download.file(url, tmp, quiet = TRUE)

lines <- readLines(tmp)
lines <- lines[2:length(lines)]
writeLines(lines)
```

* Then I use R to produce the very file you are now reading.

```r
R
rmarkdown::render("README.Rmd", output_format = "md_document")
```

## Installation

To use the software in this and related repositories (more on this below) you
need to clone the source code from [2DII's organization on
GitHub](https://github.com/2DegreesInvesting/).

The PACTA_analysis repository contains an R package called PACTA.analysis (the
name of an R package can't include "_"). Use it from inside PACTA_analysis with:

```r
devtools::load_all()`
```

## Reproducible examples

This section shows reproducible examples of both the single- and multiple-inputs
workflows. You may use these examples in at least three ways:

* Read this document for an overview of which workflows are possible.
* Run them on your local computer to track and explore their implementation.
* Run them on a server as an integration test.

### Single-inputs workflow

TODO: Document inputs/outputs

> There's a pretty important absence here... any explanation of where the critically important input files are or need to be. Hypothetically, the repo should work because it includes some default example files, but any analyst truly using this repo... the first thing they would want to do is put their input files in working_dir -- CJ

The single-inputs workflow is composed of multiple steps, each in its own
script: `r list.files(pattern = "web_tool_script")`.

With R, you can run each script individually with something like
`source("web_tool_script_1.R")` or run multiple scripts at once. For example,
you can run scripts 1 and 2 with:

```{r}
# Helper from the package PACTA.analysis
source_web_tool_scripts(1:2)
```

With the terminal, you can run each script individually. For example, you can
run script 3 with:

```{bash comment="$", error=TRUE}
Rscript --vanilla web_tool_script_3.R TestPortfolio_Input  
```

### Multiple-inputs workflow

TODO
