---
title: "Mauro's first run of the web tool"
date: "`r Sys.Date()`"
output: github_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```

Packages

```{r}
library(testthat)
library(glue)
library(config)
library(rlang)
library(here)
library(fs)
library(tidyverse)
```

Required paths

```{r}
file_name <- "TestPortfolio_Input.csv"
example_dataset <- here::here("sample_files", "20_input_files", file_name)
expected_dataset <- here::here("working_dir", "20_Raw_Inputs", file_name)
```

Required directories

```{r}
# Similar to `create_portfolio_subfolders()`
create_directory_if_it_doesnt_exist <- function(directory) {
  if (!fs::dir_exists(directory)) {
    fs::dir_create(directory)
  }
  
  invisible(directory)
}

children <- c("30_Processed_Inputs", "40_Results", "50_Outputs")
(paths <- here("working_dir", children))

walk(paths, create_directory_if_it_doesnt_exist)
```

Required repos

```bash
# Terminal
gh repo fork --clone 2DegreesInvesting/pacta-data
gh repo fork --clone 2DegreesInvesting/create_interactive_report
gh repo fork --clone 2DegreesInvesting/PACTA_analysis
```

Required branch: Run from inside the project PACTA_analysis/, particularly on the branch `current_web_functionality`.

```bash
# Terminal
cd PACTA_analysis
```

Required working directory

```{r}
here()
```

NOTE: The branch `current_web_functionaliry` is long-lived, which makes the workflow more complex than the standard GitHub workflow (by which the only long-lived branch is `master`). I suggest merging `current_web_functionality` into master, or extracting the non-overlapping code into a separate repo.

## [Instructions by Jacob](https://bit.ly/2RCRJn7)

### 1. Data 

Here we use an example dataset, which the analysis expects in a specific folder.

Sanity checks:

```{r}
stopifnot(file_exists(example_dataset))

if (file_exists(expected_dataset)) {
  warn(glue("Removing existing file: {expected_dataset}"))
  file_delete(expected_dataset)
}

file_copy(example_dataset, expected_dataset)

expect_true(file_exists(expected_dataset))
```

### 2. Assign a value to `portfolio_name_ref_all`

2. Set "TestPortfolio_Input" as the portfolio_name_ref_all in all three web_tool_scripts (lines 30, 24 and 21)

I explore what value is currently assigned to the variable `TestPortfolio_Input`:

```{r}
show_pattern_in_file <- function(file, pattern) {
  grep(pattern, readLines(file), value = TRUE)
}

(files <- dir_ls(regexp = "web_tool_script"))
this_pattern <- "portfolio_name_ref_all.*<-"
matched <- map(files, show_pattern_in_file, pattern = this_pattern)
```

It looks like the values are already the ones we expect:

```{r}
walk(matched, writeLines)

script_has_this_pattern <- grepl(this_pattern, matched)
expect_true(all(script_has_this_pattern))
```

NOTE: It seems useful to make `portfolio_name_ref_all` a parameter; that way the user can set it without changing source code.

### 3 Configuration files

3.1. Ensure this config file:

```{r}
config_1 <- here(
  "working_dir",
  "10_Parameter_File",
  "TestPortfolio_Input_PortfolioParameters.yml"
)

writeLines(readLines(config_1))

expect_true(file_exists(config_1))
```

3.1. Ensure this other configuration file also exists, and has correct paths:

```{r}
config_2 <- here("parameter_files", "WebParameters_2dii.yml")
writeLines(readLines(config_2))
```

```{r}
expect_true(file_exists(config_2))
```

```{r}
config_paths <- config::get(file = config_2)$paths
str(config_paths)

all_paths_exist <- all(map_lgl(config_paths, dir_exists))
expect_true(all_paths_exist)
```

NOTE: It would be nice to set these paths from a more flexible interface, like the arguments of a function or the parameters of a parametrized rmarkdown file.

### 4. Run scripts

* web_tool_script_1.R will produce processed input files in working_dir/30_Processed_Inputs/

```{r, error=TRUE, message=FALSE}
source("web_tool_script_1.R")
```

NOTE: Some warnings may be avoided if required directories are created only if they don't already exist.

```r
Warning message:
In read_file(paste0(file_location, "/fund_data.fst")) :
  ~/git/pacta-data/2019Q4/cleaned_files/fund_data.fst does not exist
```

NOTE: I'm not sure if this dataset is crucial, but it's missing from my clone of pacta-data/:

```{r}
dir_ls(path("..", "pacta-data", "2019Q4", "cleaned_files"))
```

* web_tool_script_2.R will take the processed inputs from the previous step, calculate PACTA results and write them to working_dir/40_Results/

NOTE: I see duplicated code in the first lines of each script. For example, you may extract all calls to `library()` into a file packages.R; and you may also extract the repeated definitions of `working_location` (which I think should be `working_location <- here::here()`).

```{r, error=TRUE, message=FALSE}
source("web_tool_script_2.R")
```

* web_tool_script_3.R this will take the results from the previous step plus some pre-calculated data from the pacta-data repo and input these results into the create_interactive_report function.

```{r}
parent <- path_dir(here())
sibling <- path(parent, "create_interactive_report")

if (!dir_exists(sibling)) {
  abort(glue("
    PACTA_analysis/ and create_interactive_report/ must share a parent directory
    See https://git.io/JUBkN
  "))
}

expect_true(dir_exists(sibling))
```

```{r, error=TRUE, cache=FALSE, message=FALSE}
source("web_tool_script_3.R") 
```

If all goes well, this will copy some of the css, js etc files into working_dir/50_Outputs/, along with the results written into the index.html file. It will also produce a zip archive of the results, which is the current not-so-pretty solution to download the report and use everything interactively when offline).

```{r}
outputs <- path("working_dir", "50_Outputs")

css <- dir_ls(outputs, recurse = TRUE, regexp = "[.]css")
expect_true(length(css) > 0L)

js <- dir_ls(outputs, recurse = TRUE, regexp = "[.]js")
expect_true(length(js) > 0L)

index <- dir_ls(outputs, recurse = TRUE, regexp = "index[.]html")
expect_true(length(index) > 0L)
```

```{r}
n <- 20L
writeLines(head(readLines(index), n))
```

```{r}
save_space <- FALSE
dir_tree(path(outputs, "TestPortfolio_Input"), recurse = save_space)
```

The output looks good!