---
output: github_document
date: "`r Sys.time()`."
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  config_general = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
devtools::load_all()
```

# PACTA_analysis

The goal of this repository is to assess how well a porfolio aligns with climate
goals. 

This documents targets internal 2DII users and developers. It provides
reproducible examples of what you can achieve with the code in this repository
and friends. You may use it as a guide to run your own analyses, or as an
integration test to ensure code changes preserve the behaviour documented here.
Readers outside 2DII may instead see other related work:
[transitionmonitor.com](https://platform.transitionmonitor.com/start),
[r2dii.data](https://github.com/2DegreesInvesting/r2dii.data),
[r2dii.match](https://github.com/2DegreesInvesting/r2dii.match), and
[r2dii.analysis](https://github.com/2DegreesInvesting/r2dii.analysis).

You may want to analyze a single portfolio and investor, or multiple ones. This
suggests two workflows:

* __Single-inputs workflow__: When the number of portfolios and investors is
exactly one, we use a specific set of scripts optimized for this purpose. This
workflow is typically used to generate interactive reports. Because this
workflow is available [online](https://platform.transitionmonitor.com/start), it
has been referred to as "the online workflow", and the scripts as "the web tool
scripts". This is misleading because this workflow can run offline too.
* __Multiple-outputs workflow__: When the number of portfolios and/or investors
is not exactly one, we use a different set of optimized scripts. This workflow
is typically used to produce static (.pdf) reports. By contrast with the
so-called "online" workflow, this one has been usually referred to as "the
offline workflow"; again, this is misleading.

Both workflows include three steps:

1. Clean portfolio: Clean the input portfolio and merge in financial data, to
categorize each holding and identify whether it's equity (EQ) or corporate bonds
(CB).
2. Run Analysis: Merge portfolios with asset-level data and scenarios, then
group results at company, portfolio and regional level.
3. Present results: Present the results in a clear output format.

--

Note that "`$`" and "`#>`" precede bash and R output, respectively.

## Setup

Let's start by loading the packages we use in this file. Notice we `load_all()`
features in the PACTA.analysis package, which lives inside the PACTA_analysis
repository.

```{r, packages}
library(devtools)
load_all()
```

```{r include=FALSE, child="helpers.Rmd"}

```

## Siblings

```{r include=FALSE}
siblings <- list.files(
  "..", pattern = "[.]pacta$", recursive = TRUE, all.files = TRUE
)
siblings <- basename(dirname(siblings))
writeLines(siblings)
```

The PACTA family includes these siblings: **`r siblings`**. They are all available
at <https://github.com/2DegreesInvesting> and some workflows assume they all
live directly under the same parent directory. Before running a workflow, you
may pull the latest commits of the relevant siblings from Github, to ensure the
code you use is up to date.

## Computing environment

TODO

```{r child="computing-environment.Rmd", eval=FALSE}

```

## Reproducible examples of PACTA workflows

This section shows reproducible examples of both the single- and multiple-inputs
workflows. You may use these examples in at least three ways:

* Read this document for an overview of which workflows are possible.
* Run them on your local computer to track and explore their implementation.
* Run them on a server as an integration test.

### Single-inputs workflow

TODO: Document inputs/outputs

> There's a pretty important absence here... any explanation of where the
critically important input files are or need to be. Hypothetically, the repo
should work because it includes some default example files, but any analyst
truly using this repo... the first thing they would want to do is put their
input files in working_dir -- CJ

The single-inputs workflow is composed of multiple steps, each in its own
script: `r list.files(pattern = "web_tool_script")`.

In R you can run each script individually with something like
`source("web_tool_script_1.R")` or run multiple scripts at once with:

```{r}
# Helper from the package PACTA.analysis
source_web_tool_scripts()
```

In a terminal you can run each script individually, e.g. "web_tool_script_1.R",
with:

```{bash comment="$", error=TRUE}
Rscript --vanilla web_tool_script_1.R TestPortfolio_Input  
```
