---
title: "Transition Disruption Metric"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
```

```{r setup}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
library(PACTA.analysis)
```

``` {r functions}
tdm_demo_plot<- function(data) {
  data %>% 
    ggplot(aes(x = year, y = production, group = type)) +
    geom_line(aes(color = type, linetype = linetype)) +
    geom_point(aes(shape = type)) +
    labs(title = "Production by Year", x = "Year", y = "Production") +
    guides(linetype = "none") +
    guides(fill = "none")
}
```

There are several metrics included in the PACTA analysis. Some metrics offer
interesting insights on short time-scales (e.g. 5 years into the future).
However, analysts may also be interested in longer-term insights (e.g. 10+ years
into the future). Here, we will propose and derive one of these metrics, called
the Transition Disruption Metric. We will then explore interesting regions
expressed by the metric, along with their interpretation.

# Introduction

PACTA offers many metrics, most of which rely heavily on the use of
forward-looking data. This data estimates the production of physical assets and
is usually able to reliably forecast data to about 5 years into the future.

Another key component in these metrics is a climate scenario, which projects 
how production *should* shift to mediate climate change. There are 
certain climate scenarios which offset the bulk of required change to the later 
years in the scenario (sometimes 20+ years into the future). So-called "delayed 
response" scenarios, such as the  Inevitable Policy Response (IPR) scenario, 
assumes that future production of most sectors follow current trends (i.e. 
business as usual), before inevitably resulting in a massive and disruptive 
transition in the long-term. Since the interesting part of these scenarios is 
often positioned well beyond the 5 year time-horizon of the data, most usual 
PACTA metrics (eg. Volume Trajectory) are ill-suited to provide useful insights. 

The purpose of these "delayed response" scenarios is to estimate the risks 
associated with a worst-case scenario. With such a scenario, we might want to 
understand how **disruptive** a climate **transition** could be to a company or 
portfolio. To help provide insights into this question, we propose the 
Transition Disruption Metric (TDM). This metric aims to provide useful insights 
in the medium- to long- term (ie. time-horizons greater than 5 years).

# Metric Definition

First, let us define (with no motivation) the Transition Disruption Metric,
$TDM$, for some technology, $i$ as:

$$TDM_i = max\left(0, \frac{S_i (t+t_2) - P_i(t+t_1)}{S_i(t+t_2) - S_i(t)} * \left(\frac{t_2}{t_2-t_1}\right)\right) $$

where:
- $i$ is some technology 

- $t$ is some year - 

- $S_i(t)$ is the production forecast by some climate scenario, for technology 
$i$, at year $t$ 

- $P_i(t)$ is the production associated to a company or portfolio, for 
technology $i$, at year $t$ 

- $t_1$ and $t_2$ are two future years, where $t_1 < t_2$

# Derivation

Now, let's derive this metric together. Consider first some arbitrary function,
$S(t)$, that in general may increase or decrease over time. This function could
represent, for example, a climate scenario, dictating how production must change
to mitigate climate change:

```{r}

data <- tibble::tribble(
  ~production, ~year,  ~type, ~linetype,
           1L, 2020L, "S(t)",   "solid",
           2L, 2025L, "S(t)",   "solid",
           3L, 2030L, "S(t)",   "solid",
           4L, 2035L, "S(t)",   "solid",
           5L, 2040L, "S(t)",   "solid",
           6L, 2045L, "S(t)",   "solid",
           7L, 2050L, "S(t)",   "solid"
  )

data %>%
  filter(type == "S(t)") %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#CC6666", "#9999CC")) +
  scale_shape_manual(values = c("circle", "triangle")) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )
```

Now, let's say we know the production associated with a company or portfolio,
$P(t)$, for the range $(t,t + t_1)$. In general, $t + t_1$ will be less than the
end year of the scenario:

```{r}

data <- tibble::tribble(
  ~production, ~year,  ~type, ~linetype,
            1, 2020L, "S(t)",   "solid",
            2, 2025L, "S(t)",   "solid",
            3, 2030L, "S(t)",   "solid",
            4, 2035L, "S(t)",   "solid",
            5, 2040L, "S(t)",   "solid",
            6, 2045L, "S(t)",   "solid",
            7, 2050L, "S(t)",   "solid",
          0.5, 2020L, "P(t)",   "solid",
            1, 2025L, "P(t)",   "solid"
  )


data %>%
  tdm_demo_plot() + 
  scale_color_manual(values = c("#9999CC", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "circle")) +
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "", "", "", "", "")
  )
```

We are now comparing projected portfolio data (accurate to the best of our
knowledge) against a scenario's production projections. For the portfolio to
"transition", it would simply need to adjust it's production to match that of
the scenario.

If we want to determine how disrupted a portfolio might be between the last date
we have data for it, $t+t_1$, and some future dae in the scenario, $t+t_2$, we
may simply draw a line connecting the two points:

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            1, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            3, 2030L,               "S(t)",   "solid",
            4, 2035L,               "S(t)",   "solid",
            5, 2040L,               "S(t)",   "solid",
            6, 2045L,               "S(t)",   "solid",
            7, 2050L,               "S(t)",   "solid",
          0.5, 2020L,               "P(t)",   "solid",
            1, 2025L,               "P(t)",   "solid",
            1, 2025L, "Required build-out",  "dashed",
            3, 2030L, "Required build-out",  "dashed"
  )


data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

Suppose now we wish to compare the slope of the scenario, $m_S$ and the slope of
the *required future build-out* of the portfolio, $m_P$ (that is, the slope of
the dashed line in the above plot).

The slope of the scenario, between $(t, t+t2)$ is: 

$$ m_S = \frac{S(t+t_2) - S(t)}{(t+t_2) - (t)} $$ 

Next let's calculate the slope of the portfolio's required build-out. Since we
have constructed this line to connect $P(t+t_1)$ and $S(t+t_2)$, we see that the
slope will be:

$$ m_P = \frac{S(t+t_2) - P(t+t_1)}{(t+t_2) - (t+t_1)}$$

Taking the ratio of these two slopes, and tidying up a bit, yields: 

$$\frac{m_P}{m_S} = \frac{S_i (t+t_2) - P_i(t+t_1)}{S_i(t+t_2) - S_i(t)} * \left(\frac{t_2}{t_2-t_1}\right)$$
This looks similar to the metric we defined above, but we aren't quite done yet.
Let's analyze some interesting regions of this function, and try to understand
thri implications.

## Interesting regions
### Increasing Scenarios
First, let's assume for now that $m_S$ is strictly positive (we will explore the
opposite case later):

- Case 1: $\frac{m_P}{m_S} > 1$: The portfolio will need to change it's
production at a rate more rapid than the scenario to achieve the same production
value at $t+t_2$.

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            1, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            3, 2030L,               "S(t)",   "solid",
            4, 2035L,               "S(t)",   "solid",
            5, 2040L,               "S(t)",   "solid",
            6, 2045L,               "S(t)",   "solid",
            7, 2050L,               "S(t)",   "solid",
          0.5, 2020L,               "P(t)",   "solid",
          0.6, 2025L,               "P(t)",   "solid",
          0.6, 2025L, "Required build-out",  "dashed",
            3, 2030L, "Required build-out",  "dashed"
  )


data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

- Case 2: $\frac{m_P}{m_S} = 1$: The portfolio and scenario have the exact same
value at $t+t_1$, and thus the portfolio will need to increase it's production
at exactly the same rate as the scenario (this is difficult to see on the graph,
since "Required build-out" is on the same line as S(t)):

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            1, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            3, 2030L,               "S(t)",   "solid",
            4, 2035L,               "S(t)",   "solid",
            5, 2040L,               "S(t)",   "solid",
            6, 2045L,               "S(t)",   "solid",
            7, 2050L,               "S(t)",   "solid",
          0.5, 2020L,               "P(t)",   "solid",
            2, 2025L,               "P(t)",   "solid",
            2, 2025L, "Required build-out",  "dashed",
            3, 2030L, "Required build-out",  "dashed"
  )

data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

- Case 3: $0 < \frac{m_P}{m_S} < 1$: The portfolio is producing more than the
scenario at $t+t_1$, but still less than is required at $t+t_2$. The slope of
required build-out will be positive, but less than the slope of the scenario.

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            1, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            3, 2030L,               "S(t)",   "solid",
            4, 2035L,               "S(t)",   "solid",
            5, 2040L,               "S(t)",   "solid",
            6, 2045L,               "S(t)",   "solid",
            7, 2050L,               "S(t)",   "solid",
          0.5, 2020L,               "P(t)",   "solid",
          2.5, 2025L,               "P(t)",   "solid",
          2.5, 2025L, "Required build-out",  "dashed",
            3, 2030L, "Required build-out",  "dashed"
  )

data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

- Case 4: $\frac{m_P}{m_S} < 0$: The portfolio is producing more than the
scenario at $t+t_2$ thus it is already more ambitious than the scenario. To
align with the scenario it would need to reduce it's production:

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            1, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            3, 2030L,               "S(t)",   "solid",
            4, 2035L,               "S(t)",   "solid",
            5, 2040L,               "S(t)",   "solid",
            6, 2045L,               "S(t)",   "solid",
            7, 2050L,               "S(t)",   "solid",
          0.5, 2020L,               "P(t)",   "solid",
            4, 2025L,               "P(t)",   "solid",
            4, 2025L, "Required build-out",  "dashed",
            3, 2030L, "Required build-out",  "dashed"
  )

data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

In this final case, we don't necessarily need a metric as we don't expect the
portfolio to be disrupted if it is already out-performing against the scenario,
so we can attenuate negative values:

$$max\left(0, \frac{m_P}{m_S}\right)$$

or 

$$max\left(0, \frac{S_i (t+t_2) - P_i(t+t_1)}{S_i(t+t_2) - S_i(t)} * \left(\frac{t_2}{t_2-t_1}\right)\right)$$
This is exactly the metric we defined above. 

### Decreasing Scenarios

To be complete, let's now consider decreasing scenarios (i.e scenarios where
$m_S < 0$). For such a scenario, the above cases still hold their validity, just
with the sign of $m_P$ reversed. That is to say:

- Case 1: The case that the portfolio needs to change more rapidly than the
scenario to meet at $t + t_2$ can still be shown as $\frac{m_P}{m_S} > 1$. This
is because, if $m_S < 0$, then $m_P$ will simply need to be more negative, and
the signs will cancel out.

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            7, 2020L,               "S(t)",    "solid",
            6, 2025L,               "S(t)",    "solid",
            5, 2030L,               "S(t)",    "solid",
            4, 2035L,               "S(t)",    "solid",
            3, 2040L,               "S(t)",    "solid",
            2, 2045L,               "S(t)",    "solid",
            1, 2050L,               "S(t)",    "solid",
            7, 2020L,               "P(t)",    "solid",
            7, 2025L,               "P(t)",    "solid",
            7, 2025L, "Required build-out",   "dashed",
            5, 2030L, "Required build-out",   "dashed"
  )

data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
    )
```

It is left as an exercise to the reader to prove that, due to similar sign
cancellations, Cases 2, 3 and 4 keep their interpretation, regardless of if the
scenario increases or decreases.

### Non-Monotonic Scenario
There is one final situation we have not yet accounted for, and this is the case
of the non-monotonic scenario. Consider the situation that $S(t)$
increases(decreases) over the period $(t, t+t_2)$, but then decreases(increases)
over the following time-period:

```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            3, 2020L,               "S(t)",    "solid",
            2, 2025L,               "S(t)",    "solid",
            1, 2030L,               "S(t)",    "solid",
            2, 2035L,               "S(t)",    "solid",
            3, 2040L,               "S(t)",    "solid",
            4, 2045L,               "S(t)",    "solid",
            5, 2050L,               "S(t)",    "solid"
  )

data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#CC6666")) +
  scale_linetype_manual(values = c("solid")) + 
  scale_shape_manual(values = c("circle")) +
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
  )
```

Here, our usual formulation might look something like this: 
```{r}

data <- tibble::tribble(
  ~production, ~year,                ~type, ~linetype,
            3, 2020L,               "S(t)",   "solid",
            2, 2025L,               "S(t)",   "solid",
            1, 2030L,               "S(t)",   "solid",
            2, 2035L,               "S(t)",   "solid",
            3, 2040L,               "S(t)",   "solid",
            4, 2045L,               "S(t)",   "solid",
            5, 2050L,               "S(t)",   "solid",
            3, 2020L,               "P(t)",   "solid",
          2.9, 2025L,               "P(t)",   "solid",
          2.9, 2025L, "Required build-out",  "dashed",
            1, 2030L, "Required build-out",  "dashed"
  )


data %>%
  tdm_demo_plot() +
  scale_color_manual(values = c("#9999CC", "#0B7A75", "#CC6666")) +
  scale_linetype_manual(values = c("dashed", "solid")) + 
  scale_shape_manual(values = c("triangle", "triangle", "circle")) +
  theme(
    axis.text.y = element_blank()
  ) + 
  scale_x_continuous(
    breaks = c(2020, 2025, 2030, 2035, 2040, 2045, 2050),
    labels = c("t", "t+t1", "t+t2", "", "", "", "")
  )
```

However this doesn't really make sense. Since, overall, the scenario is increasing, if the portfolio or company is producing more than the scenario projects, it shouldn't be disrupted at all. 

Normally, this would be accounted for by our $max(0, \frac{m_P}{m_S})$ function, however the function is now off by a factor of $-1$, since the scenario is changing directions. 

To account for this, we add a special considerations to the $TDM$ metric:

IF $S(t+t_2) > S(t) > S(t+t_1)$ OR $S(t+t_2) < S(t) < S(t+t_1)$, then: 

$$TDM = max(0, - \frac{m_P}{m_S})$$

ELSE: 

$$TDM = max(0, \frac{m_P}{m_S})$$

## TDM In Practice

One final note: in practice, PACTA's forward-looking data goes 5 years into the
future, so $t_1 = 5$, and currently we set $t_2 = 10$ by default. In this case,
this metric reduces to:

$$TDM_i = 2 * max\left(0, \frac{S_i (t+10) - P_i(t+5)}{S_i(t+10) - S_i(t)}\right) $$
